Classless Java

### 2015-11-22 paper/sections/CaseStudy.tex
section for SQL language case study

\subsection{Other case studies}
\marco{should we put here the DB with fluent setters?}

>>>>

\subsection{Embedded DSLs with Fluent Interfaces}
Since the style of fluent interfaces was invented in Smalltalk as method
cascading, more and more languages came to support fluent interfaces, including
JavaScript, Java, C++, D, Ruby, Scala, etc. For most languages, to create fluent
interfaces, programmers have to either handwritten everything or create a
wrapper around the original non-fluent interfaces, using \textbf{this}. In Java,
several libraries help create fluent APIs easily (by code generation), including
jOOQ, op4j, fluflu, JaQue, etc. These libraries have given us hints on the
implementation of fluent setters in ClassLess Java.

When creating fluent interfaces, our approach has two main advantages:
\begin{enumerate}
\item Instead of forcing programmers to hand write code using \textbf{return
    this}, our approach with \mixin annotation removes this verbosity and
  automatically generate fluent setters.
\item Along the extension direction, the return types of fluent setters are
  automatically refined.
\end{enumerate}

\noindent We use embedded DSLs of two simple SQL query languages to illustrate.
The first query language \texttt{Database} (conceptually) models
\texttt{select}, \texttt{from} and \texttt{where} clause:
\lstinputlisting[linerange=6-13]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE

\noindent Note that we are only illustrating the idea of modeling embedded DSLs
with the database example, no concrete implementation of database connection nor
operations are provided in this case study. One benefit that fluent methods give
us is method chaining convenience. So when making a query using \texttt{select},
\texttt{from} and \texttt{where}, instead of writing code like:
\lstinputlisting[linerange=26-29]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=NORMAL_QUERY

\noindent We can write more concise code below:
\lstinputlisting[linerange=35-38]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY1

\noindent Now we extend the first query language with a new feature
\texttt{orderBy} which orders the result records by a field that users
specified. The key is to extend interface \texttt{Database} with only the
feature we need to add and the return type of fluent setters in
\texttt{Database} is automatically refined to \texttt{ExtendedDatabase}:
\lstinputlisting[linerange=17-20]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE_EXT

% \lstinputlisting[linerange=44-48]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY2
